---
title: "foofactors"
author: "YanchaoLuo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction
Factors are a very useful type of variable in R, but they can also drive you nuts. Especially the "stealth factor" that you think of as character.

The goal of `foofactor` is to practice and develop an R package. The `foofactors` package has five functions:

+ `fbind`: Create a new factor from two existing factors, where the new factor's levels are the union of the levels of the input factors.

+ `freq_out`: Make a frequency table for a factor.

+ `order_appeared`: Set levels of a factor to the order appeared in data.

+ `reorder_new`: Reorder levels in descending order.

+ `detect_factor_be_character`: Detect factors be character or not.

## Installation

```{r}
# install.packages("devtools")
library(devtools)
devtools::install_github("yanchaoluo/foofactors",build_vignettes = TRUE)
library(foofactors)
```

### Binding two factors using `fbind()`:

The `fbind` function puts two factors together as one.

```{r}
factor1<- factor(c("This", "is", "test", "for"))
factor2 <- factor(c("how", "to", "use", "fbind","function"))
```

If we only use `c` function, which we can not get the expected result.

```{r}
c(factor1, factor2)
```

Using `fbind()` function.

```{r}
fbind(factor1, factor2)
```

### Make a frequency table for a factor: `freq_out()`

The `freq_out` gives a dataframe with factors and their frequency.

If we could use `as.data.frame` function.
```{r}
set.seed(1)
sample_letters <- factor(sample(letters[1:4], size = 100, replace = TRUE))
as.data.frame(table(sample_letters))
```

The `freq_out()` function is much easy to make the frequency table.

```{r}
freq_out(sample_letters)
```

### Detect factors that should be character: `detect_factor_be_character`

It's hard to figure out factor and character in R, for example:
```{r}
factor1<- class(factor(c("apple", "pear", "banana")))
factor2<- class(factor(c("apple", "pear", "apple")))
```

From above code, both "factor1" and "factor2" are defined as the factor. However, "factor1" should be a character because of the number of unique value equal to the length. 

Using `detect_factor_be_character` function can easily find "character", "factor" and neither "character" or "factor."

```{r}
factor1 <- factor(c("apple", "pear", "banana"))
detect_factor_be_character(factor1)
```

```
factor2 <-  factor(c("apple", "pear", "apple"))
factor3<- c(1,2,2,3)
detect_factor_be_character(factor2)
detect_factor_be_character(factor3) 
```
The R code above will show two types of error when you use "detect_factor_be_character."

### Reorder level of factor: 

This function is very similar with `fct_rev` function.
and `reorder_new` fuction is using desc() in dplyr to reorder factor levels in descending order.

```{r}
reorder_new(factor(c("1","2","3")))
```

From the code above, we could find that the orginal order levels is 1, 2, 3. Using the `reorder_new` function we get decending order 3, 2, 1.

### Sets levels to the order in which they appear in the data: order_appeared

Here I create a function which set the order following appearance level. and this function will only return unique input levels,

```{r}
factor(c("pear","apple","banana","banana","apple"))
order_appeared (c("pear","apple","banana","banana","apple"))
```

If we only use `factor`function, the factor level will follow the alphabet. Using `order_appeared` function, the factor level will order in which they appear in the data.
